---
title: "Practicum 1"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---
---
Author: Congkai Sun;  sun.con@northeastern.edu
        YiHsin Wen;  wen.yih@northeastern.edu
---
```{r}
library(RMySQL)

db_user <- 'root'
db_password <- 'tobenmlgb'
db_name <- 'cs5200dbsun'
db_host <- 'localhost'
db_port <- 3306

mydb <- dbConnect(MySQL(), user = db_user, password = db_password, dbname = db_name, host = db_host, port = db_port)
```

```{r}
library(tidyverse)
```
```{r}
library('plyr')
```
```{r}
library(ggplot2)
```

Problem 1
```{sql connection=mydb}
DROP TABLE IF EXISTS incidents
```

```{sql connection=mydb}
CREATE TABLE incidents(
  iid INTEGER PRIMARY KEY AUTO_INCREMENT,
  incidentDate DATE DEFAULT '0000-00-00',
  origin NUMERIC,
  airline VARCHAR(35) DEFAULT 'unknown',
  aircraft VARCHAR(35),
  flightPhase VARCHAR(35) DEFAULT 'unknown',
  impact VARCHAR(35),
  cond NUMERIC,
  FOREIGN KEY(cond) REFERENCES conditions(cid),
  FOREIGN KEY(origin) REFERENCES states(recordID)
);
```

```{sql connection=mydb}
DROP TABLE IF EXISTS states
```

```{sql connection=mydb}
CREATE TABLE states(
  recordID NUMERIC,
  PRIMARY KEY(recordID),
  FOREIGN KEY(recordID) REFERENCES airports(aid)
)
```

```{sql connection=mydb}
DROP TABLE IF EXISTS airports
```

```{sql connection=mydb}
CREATE TABLE airports(
  aid NUMERIC,
  state VARCHAR(35),
  airportName VARCHAR(35) DEFAULT 'unknown',
  airportCode CHAR(3),
  PRIMARY KEY(aid)
)
```

```{sql connection=mydb}
DROP TABLE IF EXISTS conditions
```

```{sql connection=mydb}
CREATE TABLE conditions(
  cid NUMERIC,
  cond VARCHAR(35),
  explanation TEXT,
  PRIMARY KEY(cid)
)
```

Replace climb with takeoff
```{sql connection=mydb}
UPDATE incidents 
SET flightPhase=REPLACE(flightPhase,'Climb','takeoff')
```

Replace take-off run with takeoff
```{sql connection=mydb}
UPDATE incidents 
SET flightPhase=REPLACE(flightPhase,'Take-off run','takeoff')
```

Replace landing roll with landing 
```{sql connection=mydb}
UPDATE incidents 
SET flightPhase=REPLACE(flightPhase,'Landing Roll','landing')
```

Replace decent with inflight
```{sql connection=mydb}
UPDATE incidents 
SET flightPhase=REPLACE(flightPhase,'Descent','inflight')
```

Replace taxi with takeoff
```{sql connection=mydb}
UPDATE incidents 
SET flightPhase=REPLACE(flightPhase,'Taxi','takeoff')
```

Replace approach with inflight
```{sql connection=mydb}
UPDATE incidents 
SET flightPhase=REPLACE(flightPhase,'Approach','inflight')
```

Replace parked with landing 
```{sql connection=mydb}
UPDATE incidents 
SET flightPhase=REPLACE(flightPhase,'Parked','landing')
```

Remove all military flights from the database
```{sql connection=mydb}
DELETE incidents, states, airports, conditions
FROM incidents
INNER JOIN states
INNER JOIN airports
INNER JOIN conditions
ON incidents.origin = states.recordID AND states.recordID = airports.aid AND incidents.cond = conditions.cid
WHERE airline = 'MILITARY'
```

Replace empty string with unknown
```{sql connection=mydb}
UPDATE incidents 
SET flightPhase= 'unknown' WHERE flightPhase IS NULL or flightPhase = '';
```

Replace empty string with unknown
```{sql connection=mydb}
UPDATE incidents 
SET airline= 'unknown' WHERE airline IS NULL or airline = '';
```

Replace impact value
```{sql connection=mydb}
UPDATE incidents 
SET impact = REPLACE(impact,'Caused damage','TRUE')
```

```{sql connection=mydb}
UPDATE incidents 
SET impact = REPLACE(impact,'No damage','FALSE')
```

```{sql connection=mydb}
SELECT * FROM incidents
```

```{sql connection=mydb}
SELECT * FROM airports
```

```{sql connection=mydb}
SELECT * FROM conditions
```

Problem 2
Read CSV file
```{r}
path <- "D:/Spring 2022/CS 5200/P1"
fn <- "BirdStrikesData.csv"

fileName <- paste(path, fn, sep = "/")

bevDF <- read.csv(fileName, header = TRUE, stringsAsFactors = FALSE)
```

Import the values of states table
```{r}
recordID <- bevDF$ï..Record.ID

dfstate <- data.frame(recordID)
```

```{r}
dbWriteTable(conn = mydb, "states", value = dfstate, append = TRUE, row.names = FALSE)
```

Import the values of airports table 
```{r}
aid <- bevDF$ï..Record.ID
state <- bevDF$Origin.State
airportName <-bevDF$Airport..Name

length(state) <- length(aid)
length(airportName) <- length(aid)

dfAir <- data.frame(aid, state, airportName)
```

```{r}
dbWriteTable(conn = mydb, "airports", value = dfAir, append = TRUE, row.names = FALSE)
```

Import the values of conditions table 
```{r}
cid <- bevDF$ï..Record.ID
cond <- bevDF$Conditions..Sky

length(cond) <- length(cid)
dfcon <- data.frame(cid, cond)
```

```{r}
dbWriteTable(conn = mydb, "conditions", value = dfcon, append = TRUE, row.names = FALSE)
```

Import the values of incidents table 
```{r}
incidentDate <-bevDF$FlightDate
origin <- bevDF$ï..Record.ID
airline <- bevDF$Aircraft..Airline.Operator
aircraft <- bevDF$Aircraft..Make.Model
flightPhase <- bevDF$When..Phase.of.flight
impact <- bevDF$Effect..Indicated.Damage
cond <- bevDF$ï..Record.ID

length(incidentDate) <- length(cond)
length(origin) <- length(cond)
length(airline) <- length(cond)
length(aircraft) <- length(cond)
length(flightPhase) <- length(cond)
length(impact) <- length(cond)

dfInc <- data.frame(incidentDate, origin, airline, aircraft, flightPhase, impact, cond)
```

Keep the mm/dd/yy of incident date
```{r}
df <- dfInc %>%
  mutate(incidentDate = str_remove_all(incidentDate, "0:00"))
```

```{r}
df$incidentDate <- as.Date(df$incidentDate, format = "%m/%d/%Y")
```

```{r}
dbWriteTable(conn = mydb, "incidents", value = df, append = TRUE, row.names = FALSE)
```

Problem 3
Display the parts of each table to consider about the impact of different conditions on bird strikes incidents and see which airports are prone to incidents under which specific conditions
```{sql connection=mydb}
DROP VIEW condition_factor
```

```{sql connection=mydb}
CREATE VIEW condition_factor AS
SELECT incidents.iid, incidents.incidentDate, incidents.impact, airports.airportName, conditions.cond
FROM incidents, airports, conditions
WHERE incidents.origin = airports.aid AND incidents.cond = conditions.cid
GROUP BY incidents.iid
```

```{sql connection=mydb}
SELECT * FROM condition_factor
```

Problem 4
```{sql connection=mydb}
SELECT flightPhase, COUNT(*) AS incidentsNumber
FROM incidents
Group BY flightPhase
```

Problem 5
```{sql connection=mydb}
SELECT flightPhase, COUNT(iid) AS incidentsNumber FROM incidents
GROUP BY flightPhase
HAVING COUNT(iid) >= (SELECT AVG(incident_count) FROM
            (SELECT COUNT(iid) as incident_count, flightPhase
             FROM incidents
             GROUP BY flightPhase)a);
```

Problem 6
Average number of bird strike incidents by month (across all years)
```{sql connection=mydb}
SELECT  MONTH(incidentDate) as month, ROUND(count(incidentDate)/12,0) AS count
       FROM incidents
       GROUP BY month
       ORDER BY month
```


Problem 7
Extract the incident's dates
```{r}
dates <- as.Date(df$incidentDate, "%m/%d/%Y")
```

```{r}
dates <- as.POSIXct(dates, format = "%m/%d/%Y")
```

Extract the incident's years
```{r}
years <- format(dates, format="%Y")
```

```{r}
year <- as.data.frame(years)
```

Count the frequency of incidents for each years
```{r}
incident_freq <- count(year)
```

Extract the frequency of incidents from 2005 to 2011
```{r}
target_freq <- incident_freq[6:12,]
```

```{r}
ggplot(data = target_freq, aes(x = years, y= freq)) + geom_bar(position = 'dodge', stat = 'identity') + ggtitle("The Number of Incidents from 2005 to 2011")
```

Problem 8
Stored procedure for insert airport name, aid, and state value 
```{sql connection=mydb}
CREATE DEFINER=`root`@`localhost` PROCEDURE `INSERTINCIDENTS`(IN airportName VARCHAR(35),
IN state VARCHAR(35),
IN aid NUMERIC)
BEGIN

INSERT INTO airports (airportName, aid, state)
VALUES (@airportName, @aid,  @state);

END
```

